[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.1304457.svg)](https://doi.org/10.5281/zenodo.1304457) 
[![License: AGPL v3](https://img.shields.io/badge/License-AGPL%20v3-blue.svg)](https://www.gnu.org/licenses/agpl-3.0) 
[![Travis CI](https://travis-ci.org/bnasr/phenocam.svg?branch=master)](https://travis-ci.org/bnasr/phenocam) 
[![Coverage status](https://codecov.io/gh/bnasr/phenocam/branch/master/graph/badge.svg)](https://codecov.io/github/bnasr/phenocam?branch=master) 


# Extracting PhenoCam data using phenocam R Package 

```{r setup, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

PhenoCam time-series are extracted time-series data obtained from ROI's for a given site. 
Each site has specific metadata including but not limited to how the site is set-up and where it is located, what vegetaion type is visible from the camera, and its climate regime. Each PhenoCam may have none to several ROI's per vegetation type. 

This R package is an effort to simplify data wrangling steps and finally merge them as a single dataframe.

# Installation
`phenocam` can be directly installed from the [GitHub repo](https://github.com/bnasr/phenocam):
```{r}
# installing the package from the GitHub repo
if(!require(devtools)) install.packages('devtools')
devtools::install_github('bnasr/phenocam')

# loading the package
library(phenocam)
```


# Usage
The vignette page contains useful information on how to use the R package.
```{r}
browseVignettes('phenocam')
```

Detailed guides on how to use each function has been provided in the manual page of each function.


