---
title: "PhenoCam Data Fusion"
author: "Bijan Seyednasrollah"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PhenoCam Data Fusion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


PhenoCam time-series are extracted time-series data obtained from ROI's for a given site. 
Each site has specific metadata including but not limited to how the site is set-up and where it is located, what vegetaion type is visible from the camera, and its climate regime. Each PhenoCam may have none to several ROI's per vegetation type. 

This R package is an effort to simplify data wrangling steps and finally merge them as a single dataframe.

I begin with several examples for extracting Phenocam time-series directly from the server:
- Exploring the PhenoCam metadata
- Downloading Phenocam time-series data
- Merging with other time-series

## Exploriong the PhenoCam metadata
We can obtain an up-to-date dataframe of the metadata of the entire PhenoCam network using the `getphenos()` function. The returning value would be a data.table in order to simplify further data exploration.

```{r}
library(data.table)
library(phenocam)

# obtaining the metadata as data.table
phenos <- getphenos()

head(phenos)
```

or we can just list the sites with flux data:
```{r}
phenofluxsites <- phenos[flux_data==TRUE&!is.na(flux_sitenames), .(PhenoCam=site, Flux=flux_sitenames)]
head(phenofluxsites)

#list deciduous broadleaf sites with flux tower
DB.flux <- phenos[flux_data==TRUE&primary_veg_type=='DB', site]
head(DB.flux)
```



## Downloading Phenocam timeseries data

To download the phenological time-series from the PhenoCam, we need to know the sitename, vegetation type and ROI ID. This information can be obtained from the [PhenoCam website](https://phenocam.sr.unh.edu/webcam/) or using the `getrois()` function:

```{r}
# obtaining the list of all the available ROI's on the PhenoCam server
rois <- getrois()

head(rois)
```

The `getphenoTS()` function can download a time-series and return the result as a `data.table`. For example, to obtain the time-series for [DB_1000](https://phenocam.sr.unh.edu/data/archive/dukehw/ROI/dukehw_DB_1000.html) from the __*dukehw*__ PhenoCam site:

```{r}

# to obtain the DB 1000  from dukehw
dukehw_DB_1000 <- getphenoTS(site = 'dukehw', vegType = 'DB', roiID = 1000, type = '3day')

head(dukehw_DB_1000)

dukehw_DB_1000[,date:=as.Date(date)]
dukehw_DB_1000[,plot(date, gcc_90)]

```




## Merge with other time-series such as flux data 
In a fully programmatic settings you can load the PhenoCam dataset, find the related flux data, load the flux data and merge everything together as follows:

```{r}
phenots <- getphenoTS(site = 'oregonMP', vegType = 'EN', roiID = 1000)

head(phenots)

fluxfile <- system.file('fluxnetrepo/FLX_US-Me2/FLX_US-Me2_FULLSET_DD.csv', package = 'phenocam')

fluxts <- read.csv(fluxfile, skip = 0)
fluxts[fluxts==-9999] <- NA
fluxts <- as.data.table(fluxts)
fluxts[,datetime:=as.POSIXct(as.character(TIMESTAMP), format='%Y%m%d')]
fluxts[,YYYYMMDD:=as.character(as.Date(datetime))]
fluxts[,YEAR:=year(datetime)]
fluxts[,DOY:=yday(datetime)]

head(fluxts)

```
