---
title: "Getting Started with the Phenocam API"
author: "Bijan Seyednasrollah"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with the Phenocam API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r libraries, include = FALSE, echo=FALSE}
library(data.table)
library(phenocamapi)
library(lubridate)
library(jpeg)
```



This R package is an effort to simplify data wrangling steps including downloading site data and metadata.


Each PhenoCam site has specific metadata including but not limited to how the site is set-up and where it is located, what vegetaion type is visible from the camera, and its climate regime. Each PhenoCam may have none to several ROI's per vegetation type. 'phenocamapi' is an interface to interact with the PhenoCam server to extract those data and process them in an R environment.


I begin with several examples for extracting Phenocam data directly from the server:

- Exploring the PhenoCam metadata
- Filtering the dataset using attributes
- Downloading Phenocam time-series data
- Extracting the list midday images
- Downloading midday images for a given time range
 
## Exploriong the PhenoCam metadata
We can obtain an up-to-date dataframe of the metadata of the entire PhenoCam network using the `get_phenos()` function. The returning value would be a data.table in order to simplify further data exploration.

```{r, fig.height=5, fig.width=6.5, message=FALSE}

# obtaining the phenocam site metadata from the server as data.table
phenos <- get_phenos()

# checking out the first few rows
head(phenos$site)

colnames(phenos)

# removing the sites with unkown MAT and MAP values
phenos <- phenos[!((MAT_worldclim == -9999)|(MAP_worldclim == -9999))]

# extracting the PhenoCam climate space based on the WorldClim dataset
# and plotting the sites across the climate space different vegetation type as different symbols and colors
phenos[primary_veg_type=='DB', plot(MAT_worldclim, MAP_worldclim, pch = 19, col = 'green', xlim = c(-5, 27), ylim = c(0, 4000))]
phenos[primary_veg_type=='DN', points(MAT_worldclim, MAP_worldclim, pch = 1, col = 'darkgreen')]
phenos[primary_veg_type=='EN', points(MAT_worldclim, MAP_worldclim, pch = 17, col = 'brown')]
phenos[primary_veg_type=='EB', points(MAT_worldclim, MAP_worldclim, pch = 25, col = 'orange')]
phenos[primary_veg_type=='AG', points(MAT_worldclim, MAP_worldclim, pch = 12, col = 'yellow')]
phenos[primary_veg_type=='SH', points(MAT_worldclim, MAP_worldclim, pch = 23, col = 'red')]
legend('topleft', legend = c('DB','DN', 'EN','EB','AG', 'SH'), 
       pch = c(19, 1, 17, 25, 12, 23), 
       col =  c('green', 'darkgreen', 'brown',  'orange',  'yellow',  'red' ))

```


## Filtering the dataset using attributes
For example we can just list the sites that have flux tower:

```{r, fig.height=5, fig.width=6.5, message=FALSE}
# store sites with flux_data available
phenofluxsites <- phenos[flux_data==TRUE&!is.na(flux_sitenames), .(PhenoCam=site, Flux=flux_sitenames)]

# see the first few rows
head(phenofluxsites)

#list deciduous broadleaf sites with flux tower
DB.flux <- phenos[flux_data==TRUE&primary_veg_type=='DB', site]

# see the first few rows
head(DB.flux)
```


## Downloading Phenocam timeseries data
PhenoCam time-series are extracted time-series data obtained from ROI's for a given site. 

To download the phenological time-series from the PhenoCam, we need to know the sitename, vegetation type and ROI ID. This information can be obtained from the <a href="https://phenocam.sr.unh.edu/webcam/">PhenoCam website</a> or using the `get_rois()` function:


```{r, fig.height=5, fig.width=6.5, message=FALSE}
# obtaining the list of all the available ROI's on the PhenoCam server
rois <- get_rois()

head(rois$roi_name)
colnames(rois)
```


The `get_pheno_ts()` function can download a time-series and return the result as a `data.table`. For example, to obtain the time-series for <a href="https://phenocam.sr.unh.edu/data/archive/dukehw/ROI/dukehw_DB_1000.html">DB_1000</a> from the __*dukehw*__ PhenoCam site, we can run the following code:

```{r, fig.height=5, fig.width=6.5, message=FALSE}
# to obtain the DB 1000  from dukehw
dukehw_DB_1000 <- get_pheno_ts(site = 'dukehw', vegType = 'DB', roiID = 1000, type = '3day')

colnames(dukehw_DB_1000)

dukehw_DB_1000[,date:=as.Date(date)]
dukehw_DB_1000[,plot(date, gcc_90, col = 'green', type = 'b')]
mtext('dukehw')
```


## Extracting the list midday images
While Phenocam site may have many images in a given day, a list of midday images (only one image for a day) can be useful for simple analysis.
```{r, fig.height=5, fig.width=6.5, message=FALSE}
# obtaining midday_images for dukehw
duke_middays <- get_midday_list('dukehw', direct = FALSE)

# see the first few rows
head(duke_middays)

# download a file
destfile <- tempfile(fileext = '.jpg')
download.file(duke_middays[1], destfile = destfile)

img <- try(readJPEG(destfile))
if(class(img)!='try-error'){
  par(mar= c(0,0,0,0))
  plot(0:1,0:1, type='n', axes= FALSE, xlab= '', ylab = '')
  rasterImage(img, 0, 0, 1, 1)
}
```



## Downloading midday images for a given time range
```{r, fig.height=5, fig.width=6.5, message=FALSE, eval=FALSE}
# download a subset
download_dir <- download_midday_images(site = 'dukehw', y = 2018, months = 4, download_dir = tempdir())

# list of downloaded files
duke_middays_path <- dir(download_dir, pattern = 'dukehw*', full.names = TRUE)

head(duke_middays_path)

```

While the purpose of this document is to show the basic commands for interacting with the PhenoCam API. More vignettes can be found in the package on how to merge external time-series (e.g. Flux data) with the PhenoCam time-series. 

